Untitled

### 2021-01-25 ms/manuscript.Rmd
Add dt vs environment figure

```{r load-data}
tar_load(refs)
tar_load(refs_other)
tar_load(climate)
tar_load(filmy_species)
tar_load(moorea_sites)
tar_load(phylosig)
tar_load(range_comparison)
tar_load(combined_species_means)
tar_load(t_test_results)
tar_load(env_range_dt_model_summary)
tar_load(recovery_species_means)
tar_load(filmy_habit)
tar_load(filmy_phy)
```

@@

We selected the island of Moorea, French Polynesia (17°32' S, 149°50' W) for our survey of filmy fern ecophysiology (Fig. 2).
Moorea is a small (135 km^2^; maximum elevation 1206 m) tropical oceanic island, located ca.
15 km NW of its larger sister island, Tahiti.
Despite its small size, it hosts a range of habitats including coastal strand, low elevation rainforest, and high elevation cloud forest.
A total of `r n_filmy_sp` filmy fern species are known from Moorea, including multiple terrestrial, saxicolous, and epiphytic species [@Murdock2003; @Ranker2005; @Nitta2011b; @Nitta2017; @Florence] (`r figure("filmy-habit")`).
The two major clades of filmy ferns, the hymenophylloid and trichomanoid lineages [@Pryer2001a], are represented on Moorea by `r n_hymeno_sp` and `r n_trich_sp` species, respectively.
We follow the taxonomic system of @Ebihara2006.

@@

We used microclimate data collected by @Nitta2020.
This dataset includes relative humidity (RH) and temperature measured every 15 min from `r climate_start` to `r climate_end` at the sampling sites with Hobo ProV2 dataloggers (Onset Computer Corporation, Bourne, Massachusetts, USA).
The dataloggers were outfitted with radiation shields (RS1 or RS3, Onset Computer Corporation, Bourne, Massachusetts, USA) to prevent direct contact with precipitation and solar radiation, and placed in pairs at each site (one 'epiphytic' datalogger at \circa{ }1.5 m on a tree, one 'terrestrial' datalogger at \circa{ }10 cm next to the tree).
During the study period, some dataloggers malfunctioned, presumably due to prolonged exposure to high humidity.
Any days during which one or more dataloggers failed are excluded from the dataset.
One terrestrial datalogger (Mt. Tohiea 393 m site) was lacking data for more than half of the survey period, but had data available for the same period from a preliminary survey conducted during the previous year (2012--2013), so the previous year's data for this site was used instead.
The dataset includes `r climate_n_days` days of data from `r climate_n_sites` sites (`r climate_n_loggers` dataloggers).

@@

In several species, gametophytes exceeded the maximum or minimum elevational range of sporophytes (Table 1) including *Callistopteris apiifolia* (C. Presl) Copel., *Crepidomanes minutum* var. 2 and 3, *Hymenophyllum digitatum* (Sw.) Fosberg, *Hymenophyllum javanicum* Spreng., *Hymenophyllum pallidum* (Blume) Ebihara & K.
Iwats., and *Polyphlebium borbonicum* (Bosch) Ebihara & Dubuisson.
All of these have sporophytes that are normally only found in cloud forest (\circa{ }500--1200 m), but gametophytes that are distributed to lower elevations.

@@

Mean daily temperature was strongly correlated with elevation, and dropped ca.
6 °C from the lowest surveyed site (201 m) to the highest (1206 m; Fig. 3).
Although epiphytic and terrestrial habitats were similar in temperature, epiphytic habitats tended to have lower daily mean RH, and greater daily variation in RH, than terrestrial ones (Fig. 3, Appendix S2).
This difference was greatest at low elevations, and smallest at high elevations.
One significant departure from these trends was observed at the Mt. Rotui 830 m site (indicated by arrows in Fig. 3).
This site is much more exposed than the other cloud forest sites (Fig. 2), and many of the plants on the N facing slope are stunted (J. Nitta, pers. obs.).
Daily variation in temperature and RH was much greater, and mean RH much lower, at this site than other high elevation sites.

@@

Ability to recover from desiccation was tested in sporophytes of `r n_dt_species_sporo` species and gametophytes of `r n_dt_species_gameto` species, including `r n_dt_species_both` species with both life phases (Appendix S3).
A general correlation of DT with habitat was observed in sporophytes (`r figure("filmy-dt")`).
Both terrestrial species (*Abrodictyum dentatum* (Bosch) Ebihara & K Iwats. and *C. apiifolia*) lacked even moderate DT, and failed to recover from the gentlest treatment of 2 d at --38 MPa (more stringent tests were not conducted on these species after confirming their lack of DT at this level).
Saxicolous species (*Crepidomanes kurzii* (Bedd.) Tagawa & K. Iwats., *Polyphlebium endlicherianum* (C. Presl) Ebihara & K. Iwats., and *Vandenboschia maxima* (Blume) Copel.) also had low DT; only *C. kurzii* was able to recover from desiccation up to --86 MPa for 2 d but not more.
All epiphytic species were capable of recovering from desiccation to a certain degree, but ability varied between species.
Most low elevation epiphytes could not tolerate long periods (15 d) of desiccation; only *Crepidomanes bipunctatum* (Poir.) Copel. was able to recover from desiccation longer than 2 d.
Most of the high elevation epiphytes could recover from 2 d treatments regardless of desiccation intensity (except for *H.* *pallidum* and *P.* *borbonicum*, which failed to recover from 2 d at --282 MPa), and many could withstand desiccation for 15 d.
Interestingly, *C. minutum* var. 2, *Hymenophyllum multifidum* (G. Forst.) Sw., and *Hymenophyllum polyanthos* (Sw.) Sw. could recover from desiccation at --86 or --282 MPa for 15 d, but not --38 MPa.

@@

Gametophytes of low elevation epiphytes had the highest DT (\circa{ }80% recovery), followed by high elevation epiphytes and saxicolous species (both with recovery from \circa{ }25% to 75%), with terrestrial species essentially non-tolerant (less than 20% recovery; Fig. 5).

@@

@@

**`r figure("filmy-habit")`** Examples of filmy ferns from French Polynesia.
(**a**–**d**) sporophytes; (**e**–**h**) gametophytes.
(**a**) *Crepidomanes humile*, a low elevation epiphyte. *Nitta 3392* (GH). 
(**b**) *Hymenophyllum multifidum*, a high elevation epiphyte growing in association with leafy liverworts and mosses. *Nitta 1257*
(GH). 
(**c**) *Polyphlebium endlicherianum*, a saxicolous species found at low to middle elevations. *Nitta 629* (GH). 
(**d**) *Callistopteris apiifolia*, a terrestrial species with sporophytes occurring only in humid cloud forests. *Nitta 3990* (GH).
(**e**) *C. apiifolia* (ribbon morphotype). Arrow indicates gemmae (asexual propagules) at tip of lobe. *Nitta 2565* (GH). 
(**f**) *Hymenophyllum polyanthos* (ribbon morphotype). *Nitta 3862* (GH).
(**g**) Filamentous gametophytes of an unidentified species of Hymenophyllaceae. *Nitta 4036* (GH). 
(**h**) *C. apiifolia*, showing detail of gemmifers (attachment points of gemmae, example indicated by arrow) and branched gemmae. *Nitta 3887* (GH). 
Scalebars: (**a**–**c**) 1.0 cm; (**d**) 5.0 cm; (**e**) 5.0 mm; (**f**) 2.0 mm; (**g**, **h**) 1.0 mm. 
Photographs by J. H. Nitta.

**`r figure("filmy-dt")`** Results of desiccation tolerance test on filmy fern sporophytes (**a**) and gametophytes (**b**) from Moorea, French Polynesia.
Recovery was measured by comparing maximum photochemical yield of photosystem II before desiccation treatment with values at 30 min, 24 hr, and 48 hr following desiccation at three different intensities for either 2 or 15 days (see Methods).
Recovery was not measured for all combinations of species and desiccation treatments; missing points indicate data not collected (not recovery of 0%).
Error bars indicate standard deviation.
*n* = 8 samples per species/treatment unless otherwise indicated.
Growth habit indicated in parentheses after each species: T, terrestrial; S, saxicolous; LE, low elevation epiphyte; HE, high elevation epiphyte

@@

# Set overall theme, saving default theme
default_theme <- theme_set(theme_bw(base_size = 10))

@@

**`r figure("filmy-habit")`**

@@

**`r figure("filmy-dt")`**

@@

**`r figure("filmy-dt")`** (continued)

@@

>>>>

```{r load-data}
tar_load(refs)
tar_load(refs_other)
tar_load(climate)
tar_load(filmy_species)
tar_load(moorea_sites)
tar_load(phylosig)
tar_load(range_comparison)
tar_load(combined_species_means)
tar_load(t_test_results)
tar_load(env_range_dt_model_summary)
tar_load(recovery_species_means)
tar_load(filmy_habit)
tar_load(filmy_phy)
tar_load(mean_vpd_sporo)
tar_load(mean_vpd_gameto)

@@

We selected the island of Moorea, French Polynesia (17°32' S, 149°50' W) for our survey of filmy fern ecophysiology (Fig. 2).
Moorea is a small (135 km^2^; maximum elevation 1206 m) tropical oceanic island, located ca.
15 km NW of its larger sister island, Tahiti.
Despite its small size, it hosts a range of habitats including coastal strand, low elevation rainforest, and high elevation cloud forest.
A total of `r n_filmy_sp` filmy fern species are known from Moorea, including multiple terrestrial, saxicolous, and epiphytic species [@Murdock2003; @Ranker2005; @Nitta2011b; @Nitta2017; @Florence] (`r figure("habit")`).
The two major clades of filmy ferns, the hymenophylloid and trichomanoid lineages [@Pryer2001a], are represented on Moorea by `r n_hymeno_sp` and `r n_trich_sp` species, respectively.
We follow the taxonomic system of @Ebihara2006.

@@

We used microclimate data collected by @Nitta2020.
This dataset includes relative humidity (RH) and temperature measured every 15 min from `r climate_start` to `r climate_end` at the sampling sites with Hobo ProV2 dataloggers (Onset Computer Corporation, Bourne, Massachusetts, USA).
The dataloggers were outfitted with radiation shields (RS1 or RS3, Onset Computer Corporation, Bourne, Massachusetts, USA) to prevent direct contact with precipitation and solar radiation, and placed in pairs at each site (one 'epiphytic' datalogger at \circa{ }1.5 m on a tree, one 'terrestrial' datalogger at \circa{ }10 cm next to the tree).
During the study period, some dataloggers malfunctioned, presumably due to prolonged exposure to high humidity.
Any days during which one or more dataloggers failed are excluded from the dataset.
One terrestrial datalogger (Mt. Tohiea 393 m site) was lacking data for more than half of the survey period, but had data available for the same period from a preliminary survey conducted during the previous year (2012--2013), so the previous year's data for this site was used instead.
The dataset includes `r climate_n_days` days of data from `r climate_n_sites` sites (`r climate_n_loggers` dataloggers).
Mean daily temperature on Moorea is strongly correlated with elevation, dropping \circa{ }6 °C from the lowest surveyed site (201 m) to the highest (1206 m); temperature does not differ significantly between epiphytic and terrestrial habitats [@Nitta2020].
However, epiphytic habitats tend to have lower daily mean RH, and greater daily variation in RH, than terrestrial ones.
This difference is greatest at low elevations, and smallest at high elevations.

@@

In several species, gametophytes exceeded the maximum or minimum elevational range of sporophytes  including *Callistopteris apiifolia* (C. Presl) Copel., *Crepidomanes minutum* var. 2 and 3, *Hymenophyllum digitatum* (Sw.) Fosberg, *Hymenophyllum javanicum* Spreng., *Hymenophyllum pallidum* (Blume) Ebihara & K.
Iwats., and *Polyphlebium borbonicum* (Bosch) Ebihara & Dubuisson.
All of these have sporophytes that are normally only found in cloud forest (\circa{ }500--1200 m), but gametophytes that are distributed to lower elevations.

@@

@@

Ability to recover from desiccation was tested in sporophytes of `r n_dt_species_sporo` species and gametophytes of `r n_dt_species_gameto` species, including `r n_dt_species_both` species with both life phases (Appendix S3).
A general correlation of DT with habitat was observed in sporophytes (`r figure("dt")`**a**).
Both terrestrial species (*Abrodictyum dentatum* (Bosch) Ebihara & K Iwats. and *C. apiifolia*) lacked even moderate DT, and failed to recover from the gentlest treatment of 2 d at --38 MPa (more stringent tests were not conducted on *A. dentatum* after confirming its lack of DT at this level).
Saxicolous species (*Crepidomanes kurzii* (Bedd.) Tagawa & K. Iwats., *Polyphlebium endlicherianum* (C. Presl) Ebihara & K. Iwats., and *Vandenboschia maxima* (Blume) Copel.) also had low DT; only *C. kurzii* was able to recover from desiccation up to --86 MPa for 2 d but not more.
All epiphytic species were capable of recovering from desiccation to a certain degree, but ability varied between species.
Most low elevation epiphytes could not tolerate long periods (15 d) of desiccation; only *Crepidomanes bipunctatum* (Poir.) Copel. was able to recover from desiccation longer than 2 d.
Most of the high elevation epiphytes could recover from 2 d treatments regardless of desiccation intensity (except for *H.* *pallidum* and *P.* *borbonicum*, which failed to recover from 2 d at --282 MPa), and many could withstand desiccation for 15 d.
Interestingly, *C. minutum* var. 2, *Hymenophyllum multifidum* (G. Forst.) Sw., and *Hymenophyllum polyanthos* (Sw.) Sw. could recover from desiccation at --86 or --282 MPa for 15 d, but not --38 MPa.

Gametophytes of low elevation epiphytes had the highest DT (\circa{ }80% recovery), followed by high elevation epiphytes and saxicolous species (both with recovery from \circa{ }25% to 75%), with terrestrial species essentially non-tolerant (less than 20% recovery; `r figure("dt")`**b**).

@@

@@

## Tables

**`r table("glmm")`**

@@

**`r figure("habit")`** Examples of filmy ferns from French Polynesia.
(**a**–**d**) sporophytes; (**e**–**h**) gametophytes.
(**a**) *Crepidomanes humile*, a low elevation epiphyte. *Nitta 3392* (GH). 
(**b**) *Hymenophyllum multifidum*, a high elevation epiphyte growing in association with leafy liverworts and mosses. *Nitta 1257*
(GH). 
(**c**) *Polyphlebium endlicherianum*, a saxicolous species found at low to middle elevations. *Nitta 629* (GH). 
(**d**) *Callistopteris apiifolia*, a terrestrial species with sporophytes occurring only in humid cloud forests. *Nitta 3990* (GH).
(**e**) *C. apiifolia* (ribbon morphotype). Arrow indicates gemmae (asexual propagules) at tip of lobe. *Nitta 2565* (GH). 
(**f**) *Hymenophyllum polyanthos* (ribbon morphotype). *Nitta 3862* (GH).
(**g**) Filamentous gametophytes of an unidentified species of Hymenophyllaceae. *Nitta 4036* (GH). 
(**h**) *C. apiifolia*, showing detail of gemmifers (attachment points of gemmae, example indicated by arrow) and branched gemmae. *Nitta 3887* (GH). 
Scalebars: (**a**–**c**) 1.0 cm; (**d**) 5.0 cm; (**e**) 5.0 mm; (**f**) 2.0 mm; (**g**, **h**) 1.0 mm. 
Photographs by J. H. Nitta.

**`r figure("dt")`** Desiccation tolerance of filmy fern sporophytes (**a**) and gametophytes (**b**) from Moorea, French Polynesia.
Recovery was measured by comparing maximum photochemical yield of photosystem II before desiccation treatment with values at 30 min, 24 hr, and 48 hr following desiccation at three different intensities for either 2 or 15 days (see Methods).
Recovery was not measured for all combinations of species and desiccation treatments; missing points indicate data not collected (not recovery of 0%).
Error bars indicate standard deviation.
*n* = 8 samples per species/treatment unless otherwise indicated.
Growth habit indicated in parentheses after each species: T, terrestrial; S, saxicolous; LE, low elevation epiphyte; HE, high elevation epiphyte

**`r figure("combined")`** Phylogeny of filmy ferns from Moorea, French Polynesia with comparison of physiological parameters (mean ± 1 SD) between sporophytes (dark grey bars) and gametophytes (light grey bars).
Phylogeny adapted from Nitta et al. (2016); only species with physiological data shown.
Bootstrap support >50% from 100 ML bootstraps indicated at nodes.
Bold letters indicate major clades (H = hymenophylloid clade, T = trichomanoid clade).
Growth habit indicated by colored symbols (blue = low elevation epiphyte, green = high elevation epiphyte, yellow = saxicolous, red = terrestrial).
**Crepidomanes bipunctatum** and **C. humile** were observed growing as both epiphytes and saxicoles, but are shown as epiphytes to distinguish them from exclusively saxicolous species.
Asterisk indicates significant difference in means (**P** <0.05) between generations for a particular species (**t**-test); ns indicates no significant difference; lack of ns or asterisk means not enough observations available for that species for **t**-test.
Left bar plot shows recovery (%) of chlorophyll fluorescence (Fv/Fm) following 2 day desiccation at –86 MPa (*Abrodictyum dentatum* was desiccated at –38 MPa instead of –86 MPa). 
Right bar plot shows maximum relative electron transport rate of photosystem II (ETR~max~). 
The other photosynthetic parameter measured, photosynthetic photon flux density at 95% saturation of RETR (PPFD~95%~), was qualitatively very similar to ETR~max~ and not shown.
Absence of a bar indicates measurements not available for gametophytes or sporophytes of that species.

**`r figure("corr")`** Recovery of F~v~/F~m~ (%) following 2 d desiccation at –86 MPa plotted against environmental vapor pressure deficit (VPD; kPa) (**a**, **b**) or range of gametophytes beyond sporophytes (m) (**c**) for gametophytes (**a**, **c**) or sporophytes (**b**). 
Species means ± 1 SD are shown. 
Growth habit indicated by colored symbols (blue = low elevation epiphyte, green = high elevation epiphyte, yellow = saxicolous, red = terrestrial). 
**Crepidomanes bipunctatum** and **C. humile** were observed growing as both epiphytes and saxicoles, but were treated as epiphytes to distinguish them from exclusively saxicolous species.
**Abrodictyum dentatum** sporophytes were desiccated at –38 MPa instead of –86 MPa.
Range of gametophytes beyond sporophytes calculated as the total elevation that gametophytes exceeded sporophytes at either end of the elevational gradient.
Relationships between recovery and VPD or range of gametophytes beyond sporophytes were analyzed with phylogenetic generalized least squares (PGLS); no significant relationships were detected.

@@

# Set overall theme, saving default theme
default_theme <- theme_set(
  theme_bw(base_size = 10) +
    theme(
      panel.grid.minor = element_blank()
    ))

@@

**`r figure("habit")`**

@@

**`r figure("dt")`**

@@

**`r figure("dt")`** (continued)

@@

```{r corr-plot, fig.width = 6.85, fig.height = 5, fig.cap = ""}
# Combine recovery, VPD, and growth habit into a single dataframe (wide format)
recovery_mean_data <- select(combined_species_means, species, generation, recovery_mean) %>%
  pivot_wider(values_from = recovery_mean, names_from = generation) %>%
  rename_with(~glue("recovery_{.}_mean"), .cols = -species)

recovery_sd_data <- select(combined_species_means, species, generation, recovery_sd) %>%
  pivot_wider(values_from = recovery_sd, names_from = generation) %>%
  rename_with(~glue("recovery_{.}_sd"), .cols = -species)

plot_data <-
list(filmy_habit, recovery_mean_data, recovery_sd_data, mean_vpd_sporo, mean_vpd_gameto, range_comparison) %>%
  reduce(left_join, by = "species")

a <- ggplot(plot_data, aes(x = vpd_sporophyte_mean, y = recovery_sporophyte_mean, color = habit)) +
  geom_linerange(aes(
    ymin = recovery_sporophyte_mean - recovery_sporophyte_sd,
    ymax = recovery_sporophyte_mean + recovery_sporophyte_sd),
    alpha = 0.3) +
  geom_linerange(aes(
    xmin = vpd_sporophyte_mean - vpd_sporophyte_sd,
    xmax = vpd_sporophyte_mean + vpd_sporophyte_sd),
    alpha = 0.3) +
  geom_point() +
  scale_y_continuous(
    labels = percent, 
    limits = c(0,1.25)) +
  scale_color_manual(values = habit_colors, labels = pretty_text) +
  labs(
    y = "Recovery",
    x = "VPD (kPa)",
    color = "Growth habit")

b <- ggplot(plot_data, aes(x = vpd_gametophyte_mean, y = recovery_gametophyte_mean, color = habit)) +
  geom_linerange(aes(
    ymin = recovery_gametophyte_mean - recovery_gametophyte_sd,
    ymax = recovery_gametophyte_mean + recovery_gametophyte_sd),
    alpha = 0.3) +
  geom_linerange(aes(
    xmin = vpd_gametophyte_mean - vpd_gametophyte_sd,
    xmax = vpd_gametophyte_mean + vpd_gametophyte_sd),
    alpha = 0.3) +
  geom_point() +
  scale_y_continuous(
    labels = percent, 
    limits = c(0,1.25)) +
  scale_color_manual(values = habit_colors, labels = pretty_text) +
  labs(
    y = "Recovery",
    x = "VPD (kPa)",
    color = "Growth habit") +
  guides(color = "none")

c <- ggplot(plot_data, aes(x = gameto_beyond_sporo, y = recovery_gametophyte_mean, color = habit)) +
  geom_linerange(aes(
    ymin = recovery_gametophyte_mean - recovery_gametophyte_sd,
    ymax = recovery_gametophyte_mean + recovery_gametophyte_sd),
    alpha = 0.3) +
  geom_point() +
  scale_color_manual(values = habit_colors, labels = pretty_text) +
  scale_y_continuous(labels = percent) +
  labs(
    y = "Recovery",
    x = "Elevation (m)",
    color = "Growth habit") +
  guides(color = "none")

a + b + c + 
  guide_area() + 
  plot_layout(guides = "collect") +
  plot_annotation(tag_levels = "a")  &
  theme(plot.tag = element_text(face = 'bold'))

ggsave(here::here("results/ms/fig_corr.pdf"), width = 6.85, height = 5)
```

**`r figure("corr")`**
